---
knitr:
  opts_chunk: 
    collapse: false
    comment: "#>" 
    fig.height: 8 
    fig.width: 8
    cache: true
---


# Map types

`mf_map()` displays map layers on a georeferenced plot. It has three main arguments:

  - `x`, an `sf` object
  -  `var`, the name(s) of a variable(s) to map
  -  `type`, the map layer type


## Example dataset

The following code imports relevant layers from the [geopackage](https://www.geopackage.org/) file **lot46.gpkg**.

```{r}
library(sf)
# import municipalities
com <- st_read("data/lot46.gpkg", layer = "commune", quiet = TRUE)
# import departments
dep <- st_read("data/lot46.gpkg", layer = "departement", quiet = TRUE)
# import restaurants
resto <- st_read("data/lot46.gpkg", layer = "restaurant", quiet = TRUE)
# import roads
route <- st_read("data/lot46.gpkg", layer = "route", quiet = TRUE)
```


## Plot a base map

`mf_map(x, type = "base")` displays `sf` objects geometries. 

```{r mf_base}
library(mapsf)
# Municipalities
mf_map(x = com, border = "white")
mf_map(x = dep, lwd = 2, col = NA, add = TRUE)
mf_map(x = route, lwd = .5, col = "ivory4", add = TRUE)
mf_map(x = resto, pch = 20, cex = .7, col = "darkred", add = TRUE)
```


## Proportional Symbols

`mf_map(x, var, type = "prop")` displays symbols with areas proportional to a quantitative variable (stocks). The `inches` argument is used to customize the symbols sizes.

```{r proportional_symbols}
# Municipalities
mf_map(x = com)
# Proportional Symbols
mf_map(
  x = com, 
  var = "POPULATION",
  type = "prop",
  leg_title = "Total population\n(2015)"
)
# Title
mf_title("Population distribution in the Lot department")
```


## Choropleth Map

In choropleth maps, areas are shaded according to the variation of a quantitative variable. They are used to represent ratios or indices.  
`mf_map(x, var, type = "choro")` displays choropleth maps . Arguments `nbreaks`, and `breaks` allow to customize the variable classification. `mf_get_breaks()` allows to classify data outside of the function itself. 



```{r choro}
# Population density
com$DENS <- 1e6 * com$POP / as.numeric(st_area(com))
mf_map(
  x = com,
  var = "DENS",
  type = "choro",
  breaks = "quantile",
  pal = "Dark Mint",
  lwd = 1,
  leg_title = "Population density\n(inh. per km2)", 
  leg_val_rnd = 0
)
mf_title("Population distribution in the Lot department")
```



### Color palettes {#palettes}

Colors palettes, defined with `pal`, can be created with `mf_get_pal()` or you can use palette names from `hcl.pals()`.

```{r pal1}
mf_map(x = com, var = "DENS", type = "choro",
       breaks = "quantile", nbreaks = 4, pal = "Mako")
```

![](img/swatch-plot-1.svg)

`mf_get_pal()` builds sequential, diverging and qualitative color palettes. Diverging color palettes can be dissymmetric (different number of colors in each of the two gradients).

```{r pal2,  fig.height=3, nm=TRUE}
mypal <- mf_get_pal(n = c(6,3), palette = c("Burg", "Teal"))
image(1:9, 1, as.matrix(1:9), col=mypal, xlab = "", ylab = "", xaxt = "n",
      yaxt = "n",bty = "n")
```




## Typology Map

`mf_map(x, var, type = "typo")` displays a typology map of a qualitative variable. 

```{r typo_simple}
mf_map(
  x = com,
  var = "STATUT",
  type = "typo",
  pal = c("aquamarine4", "yellow3","wheat"),
  lwd = .7,
  leg_title = ""
)
mf_title("Municipality status")
```


`val_order` can be used to set the modalities order in the legend.

```{r typo_order}
mf_map(
  x = com,
  var="STATUT",
  type = "typo",
  pal = c("aquamarine4", "yellow3","wheat"),
  val_order = c("Préfecture", "Sous-préfecture", "Commune simple"),
  lwd = .7,
  leg_title = ""
)
mf_title("Municipality status")
```



## Proportional Symbols using Choropleth Coloration

`mf_map(x, var, type = "prop_choro")` creates a map of symbols that are proportional to values of a first variable and colored to reflect the classification of a second variable.


```{r choroprop}
mf_map(x = com)
mf_map(
  x = com,
  var = c("POPULATION", "DENS"),
  type = "prop_choro",
  border = "grey50",
  lwd = 1,
  leg_pos = c("topright", "topleft"),
  leg_title = c("Population", "Population density\n(inh. per km2)"), 
  breaks = "q6",
  pal = "Magenta",
  leg_val_rnd = c(0,1))
mf_title("Population")
```

## Proportional Symbols using Typology Coloration

`mf_map(x, var, type = "prop_typo")` creates a map of symbols that are proportional to values of a first variable and colored to reflect the modalities of a second qualitative variable.

```{r typoprop}
mf_map(x = com)
mf_map(
  x = com,
  var = c("POPULATION", "STATUT"),
  type = "prop_typo",
  pal = c("aquamarine4", "yellow3","wheat"),
  val_order = c("Préfecture", "Sous-préfecture", "Commune simple"),
  leg_pos = c("bottomleft","topleft"),
  leg_title = c("Population\n(2015)",
                "Municipality status"),
)
mf_title("Population")
```
  

